---
title: "Uncertainty Representation: Experiment 2 Analysis"
author(s): "Joey Heffner"; "Marc Lluis Vives"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

# Libraries

You will need the following libraries to run everything in the script. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, message = FALSE)

# Libraries
library(here)         # relative paths
library(janitor)      # clean names
library(tidyverse)    # tidy functions
library(broom)        # convert stats to tidy
library(knitr)        # knit functions
library(kableExtra)   # extra markdown functions
library(ggrepel)      # geom_text_repel
library(gridExtra)    # aesthetics 
library(purrr)        # map functions
library(rjson)        # json reader
library(jsonlite)     # json reader
library(abind)        # abind
library(multiway)     # INDSCAL
library(corrplot)     # correlation plots
library(lme4)         # mixed effects
library(lmerTest)     # mixed effects 
library(Matrix)       # matrix manipulation functions
library(jmuOutlier)   # permutation tests
library(ggdendro)     # hc in ggplot
library(latex2exp)    # latex expressions
```

# Load data 

```{r load, message=FALSE, warning=FALSE} 
# Relative paths
dir_project <- here()
dir_study <- str_c(dir_project, "/experiment2")
dir_clean <- str_c(dir_study, "/data/clean")
dir_graphs <- str_c(dir_study, "/graphs")

# Data
df_indiv_measures <-  read_csv(str_c(dir_clean, "/individual_measures.csv"))
df_darm <- read_csv(str_c(dir_clean, "/darm_uncertainty.csv"))
df_gambles <- read_csv(str_c(dir_clean, "/gambles.csv"))
df_internal_external <- read_csv(str_c(dir_clean, "/internal_external.csv"))
df_spam <- read_csv(str_c(dir_clean, "/meadows_spam_simple.csv"))
df_prob <- read_csv(str_c(dir_clean, "/probability.csv"))
df_gamble_ids <- read_csv(str_c(dir_clean, "/gamble_ids.csv"))
```

# Figure 1 - Methods

See publication for figure 1. 

# Figure 2 - Experiment 1

See experiment1_analysis script.

# Figure 3 - Conceptual representations of uncertainty are one dimensional and organized by valence or probability

A) The percentage of variance explained for principal dimension from the similarity data reveals a sharp kink after the first component, illustrating that the variance is best explained by one component.
B) Plot of solution of multidimensional scaling of the similarity. data using the first two dimensions. 
C) Group-level representational dissimilarity matrix of uncertainty words, organized by hierarchical cluster.

```{r fig3s}
## Figure 3 Group RDM
df_to_matrix <- function(df) {
  
  unc_matrix <- matrix(df$mean_dist, nrow = 18, ncol = 18, byrow=TRUE, 
                       dimnames = list(c(unique(df$uncertain1)), c(unique(df$uncertain2))))
  
  return(unc_matrix)
}

fig3_group <- df_spam %>% 
  group_by(uncertain1, uncertain2) %>% 
  summarise(mean_dist = mean(rdm_values))
fig3_group_rdm <- df_to_matrix(fig3_group)

fig3_mds <- cmdscale(fig3_group_rdm, eig=TRUE)
fig3_eigens <- fig3_mds$eig
pca1_var <- fig3_eigens[1] / sum(fig3_eigens)
pca2_var <- fig3_eigens[2] / sum(fig3_eigens)
pca3_var <- fig3_eigens[3] / sum(fig3_eigens)

## Figure 3a
fig3a_data <- data.frame(var = fig3_mds$eig / sum(fig3_mds$eig), 
                         pca = seq(1, length(fig3_mds$eig)))

fig3a_plot <- ggplot(fig3a_data, aes(x = pca, y = var)) + 
  geom_point(size = 2) + 
  geom_line() + 
  scale_x_continuous(name = "Principal Component", breaks = c(5, 10, 15)) + 
  scale_y_continuous(name = "Variance Explained (%)", label = scales::percent) + 
  theme_classic() + 
  theme(text = element_text(size = 10))
fig3a_plot
ggsave(filename=str_c(dir_graphs, "/fig3a.pdf"), fig3a_plot , width = 3, height = 2.5, useDingbats=F)

## Figure 3b
fig3b_data <- as.data.frame(fig3_mds$points) %>%
  rename(pca1 = V1, pca2 = V2) %>%
  rownames_to_column(var = "uncertain")

fig3b_plot <- ggplot(fig3b_data, aes(x = pca1, y = pca2)) + 
  geom_point() + 
  geom_text_repel(aes(label = uncertain)) + 
  xlab(str_c("PCA1: ", as.character(round(pca1_var*100)), "%")) + 
  ylab(str_c("PCA2: ", as.character(round(pca2_var*100)), "%")) + 
  theme_classic() + 
  theme(text = element_text(size = 12), 
        aspect.ratio = 1)
fig3b_plot
ggsave(filename=str_c(dir_graphs, "/fig3b.pdf"), fig3b_plot , width = 4, height = 4, useDingbats=F)


## Figure 3c - RDM with clustering

# Hierarchical clustering 
fig3c_hc <- hclust(as.dist(fig3_group_rdm), method = "complete")

cluster_cols <- as.data.frame(cutree(fig3c_hc, k = 2)) %>%
  rownames_to_column(var = "uncertain") %>%
  rename(clusters = `cutree(fig3c_hc, k = 2)`)

fig3c_data <- fig3_group %>%
  left_join(., cluster_cols %>% select(cluster1 = clusters, uncertain1 = uncertain), by = "uncertain1") %>%
  left_join(., cluster_cols %>% select(cluster2 = clusters, uncertain2 = uncertain), by = "uncertain2")

fig3c_plot <- ggplot(fig3c_data, aes(x = reorder(uncertain1, cluster1), y = reorder(uncertain2, cluster2))) + 
  geom_tile(aes(fill = mean_dist)) + 
  scale_fill_gradient2(low = "red", mid = "orange", midpoint = mean(fig3c_data$mean_dist), high = "yellow") + 
  theme_classic() + 
  theme(text = element_text(size = 16), 
        axis.text.x = element_text(angle = 45, hjust = 1), 
        aspect.ratio = 1)
fig3c_plot
ggsave(filename=str_c(dir_graphs, "/fig3c.pdf"), fig3c_plot , width = 6, height = 6, useDingbats=F)

## Correlation MDS solution x valence
fig3e_valence <- df_darm %>% 
  group_by(uncertain_word) %>%
  summarise(mean_v = mean(valence), sd_v = sd(valence), 
            n_v = n(), se_v = sd_v / sqrt(n_v))

fig3e_data <- as.data.frame(fig3_mds$points) %>%
  rename(pca1 = V1, pca2 = V2) %>%
  #rownames_to_column(var = "uncertain_word") %>%
  cbind(fig3e_valence)

fig3e_cor <- cor.test(fig3e_data$pca1, fig3e_data$mean_v, method = "spearman")


## MDS solution x probability 
fig3f_prob <- df_prob %>% 
  group_by(uncertain_word) %>%
  summarise(mean_p = mean(probability/100), sd_p = sd(probability/100), 
            n_p = n(), se_p = sd_p / sqrt(n_p))

fig3f_data <- as.data.frame(fig3_mds$points) %>%
  rename(pca1 = V1, pca2 = V2) %>%
  #rownames_to_column(var = "uncertain_word") %>%
  cbind(fig3f_prob)

fig3f_cor <- cor.test(fig3f_data$pca1, fig3f_data$mean_p, method = "spearman")


## Arousal
fig3g_arousal <- df_darm %>% 
  group_by(uncertain_word) %>%
  summarise(mean_a = mean(arousal), sd_a = sd(arousal), 
            n_a = n(), se_a = sd_a / sqrt(n_a))

fig3g_data <- as.data.frame(fig3_mds$points) %>%
  rename(pca1 = V1, pca2 = V2) %>%
  #rownames_to_column(var = "uncertain_word") %>%
  cbind(fig3g_arousal)
fig3g_cor <- cor.test(fig3g_data$pca1, fig3g_data$mean_a, method = "spearman")


## Internal externa;
fig3h_internal_external <- df_internal_external %>% 
  group_by(uncertain_word) %>%
  summarise(mean_i = mean(internal_external), sd_i = sd(internal_external), 
            n_i = n(), se_i = sd_i / sqrt(n_i))

fig3h_data <- as.data.frame(fig3_mds$points) %>%
  rename(pca1 = V1, pca2 = V2) %>%
  #rownames_to_column(var = "uncertain_word") %>%
  cbind(fig3h_internal_external)
fig3h_cor <- cor.test(fig3h_data$pca1, fig3h_data$mean_i, method = "spearman")

# All correlations
fig3e_cor # valence     r = -0.71
fig3f_cor # probability r = -0.90
fig3g_cor # arousal     r = -0.37
fig3h_cor # int-ext     r = 0.30


## Correlation between prob and valence
fig3g_data <- fig3e_valence %>%
  cbind(fig3f_prob)

cor.test(fig3g_data$mean_v, fig3g_data$mean_p, method = "spearman")
```

# Intext analysis - Probability and valence contribute to structuring uncertainty representations

```{r intext_3.2}
##  Correlation of gambling behavior and of spam (and all others)

#### Gamble matrix ####
fig3a_gamble <- df_gambles %>%
  filter(sub != 59, sub != 115) %>% # excluded for having 0 variance in other tasks (can't correlate)
  mutate(choice = case_when(choice == 2 ~ 1, 
                            choice == 1 ~ 0)) %>%
  arrange(sub, gamble)

n_subs <- fig3a_gamble %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig3a_gamble %>% select(sub) %>% unique() %>% pull()
n_gambles <- fig3a_gamble %>% select(gamble) %>% unique() %>% tally() %>% pull()

fig3a_gamble_matrix <- matrix(1, n_gambles, n_subs)

for (i in seq(1:n_subs)) {
  fig3a_gamble_matrix[,i] <- fig3a_gamble %>% filter(sub == sub_nums[i]) %>% pull(choice)
}

#### Valence & arousal matrices ####
fig3a_darm <- df_darm %>%
  filter(sub != 59, sub != 115) %>%
  arrange(sub, uncertain_word)

n_subs <- fig3a_darm %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig3a_darm %>% select(sub) %>% unique() %>% pull()
n_words <- fig3a_darm %>% select(uncertain_word) %>% unique() %>% tally() %>% pull()

fig3a_valence_matrix <- matrix(1, n_words, n_subs)
fig3a_arousal_matrix <- matrix(1, n_words, n_subs)

for (i in seq(1:n_subs)) {
  fig3a_valence_matrix[,i] <- fig3a_darm %>% filter(sub == sub_nums[i]) %>% pull(valence)
  fig3a_arousal_matrix[,i] <- fig3a_darm %>% filter(sub == sub_nums[i]) %>% pull(arousal)
}

#### Probability matrix ####
fig3a_prob <- df_prob %>%
  filter(sub != 59, sub != 115) %>%
  arrange(sub, uncertain_word)

n_subs <- fig3a_prob %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig3a_prob %>% select(sub) %>% unique() %>% pull()
n_words <- fig3a_prob %>% select(uncertain_word) %>% unique() %>% tally() %>% pull()

fig3a_prob_matrix <- matrix(1, n_words, n_subs)

for (i in seq(1:n_subs)) {
  fig3a_prob_matrix[,i] <- fig3a_prob %>% filter(sub == sub_nums[i]) %>% pull(probability)
}

#### Internal external matrix ####
fig3a_intext <- df_internal_external %>%
  filter(sub != 59, sub != 115) %>%
  arrange(sub, uncertain_word)

n_subs <- fig3a_intext %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig3a_intext %>% select(sub) %>% unique() %>% pull()
n_words <- fig3a_intext %>% select(uncertain_word) %>% unique() %>% tally() %>% pull()

fig3a_intext_matrix <- matrix(1, n_words, n_subs)

for (i in seq(1:n_subs)) {
  fig3a_intext_matrix[,i] <- fig3a_intext %>% filter(sub == sub_nums[i]) %>% pull(internal_external)
}

#### SPAM matrix ####
fig3a_spam <- df_spam %>%
  filter(sub != 59, sub != 115) %>%
  arrange(sub, uncertain1, uncertain2)

n_subs <- fig3a_spam %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig3a_spam %>% select(sub) %>% unique() %>% pull()
n_words <- fig3a_spam %>% select(uncertain1, uncertain2) %>% unique() %>% tally() %>% pull()

fig3a_spam_matrix <- matrix(1, n_words, n_subs)

for (i in seq(1:n_subs)) {
  fig3a_spam_matrix[,i] <- fig3a_spam %>% filter(sub == sub_nums[i]) %>% pull(rdm_values)
}

#### Create correlation matrices
fig3a_gamble_cor_matrix <- cor(fig3a_gamble_matrix, method = "spearman") %>% as.matrix()
fig3a_valence_cor_matrix <- cor(fig3a_valence_matrix, method = "spearman") %>% as.matrix()
fig3a_arousal_cor_matrix <- cor(fig3a_arousal_matrix, method = "spearman") %>% as.matrix()
fig3a_intext_cor_matrix <- cor(fig3a_intext_matrix, method = "spearman") %>% as.matrix()
fig3a_prob_cor_matrix <- cor(fig3a_prob_matrix, method = "spearman") %>% as.matrix()
fig3a_spam_cor_matrix <- cor(fig3a_spam_matrix, method = "spearman") %>% as.matrix()

## fig3a permutation testing
# Goal: Create 100,000 random permutations of both gamble and valence behaviors 
#       by shuffling participant data and doing a correlation. Store the spearman
#       correlation and repeat

lead_wrap <- function(y, n) {
  # Function wraps vector on itself according to a lead
  part1 <- lead(y, n-1) # goes from n-1 to end of y
  part2 <- y[1:n-1] # want to go from beginning of y to n-1
  
  # Remove NA
  y_new <- c(part1[!is.na(part1)], part2)
  return(y_new)
}

lag_wrap <- function(y, n) {
  # Function wraps vector on itself according to a lead
  part1 <- lag(y, n-1) # shifts beginning of y down n-1
  part2 <- tail(y, n-1) # want to get last n-1 elements
  
  # Remove NA
  y_new <- c(part2, part1[!is.na(part1)])
  return(y_new)
}

num_perm <- 100000
gamble_valence_r <- data.frame(r = vector(length = num_perm), p = vector(length = num_perm))
gamble_spam_r <-  data.frame(r = vector(length = num_perm), p = vector(length = num_perm))
gamble_arousal_r <-  data.frame(r = vector(length = num_perm), p = vector(length = num_perm))
gamble_prob_r <-  data.frame(r = vector(length = num_perm), p = vector(length = num_perm))
gamble_intext_r <-  data.frame(r = vector(length = num_perm), p = vector(length = num_perm))

for (j in seq(1:num_perm)) {
  print(str_c(as.character(round(j/num_perm, 5)*100), "%"))
  
  # Permute cor matrix of all except gamble_cor_matrix
  x <- sample(dim(fig3a_valence_cor_matrix)[1])
  
  # Remake cor matrix with ordering using x
  fig3a_spam_cor_matrix_permute <- matrix(1, n_subs, n_subs)
  fig3a_valence_cor_matrix_permute <- matrix(1, n_subs, n_subs)
  fig3a_arousal_cor_matrix_permute <- matrix(1, n_subs, n_subs)
  fig3a_intext_cor_matrix_permute <- matrix(1, n_subs, n_subs)
  fig3a_prob_cor_matrix_permute <- matrix(1, n_subs, n_subs)
  
  for (i in seq(1:length(x))) {
    # make everything 1's at the top
    fig3a_valence_cor_matrix_permute[,i] <- lead_wrap(y = fig3a_valence_cor_matrix[,x[i]], n = x[i])
    fig3a_arousal_cor_matrix_permute[,i] <- lead_wrap(y = fig3a_arousal_cor_matrix[,x[i]], n = x[i])
    fig3a_spam_cor_matrix_permute[,i] <- lead_wrap(y = fig3a_spam_cor_matrix[,x[i]], n = x[i])
    fig3a_prob_cor_matrix_permute[,i] <- lead_wrap(y = fig3a_prob_cor_matrix[,x[i]], n = x[i])
    fig3a_intext_cor_matrix_permute[,i] <- lead_wrap(y = fig3a_intext_cor_matrix[,x[i]], n = x[i])
    
    # put into right place
    fig3a_valence_cor_matrix_permute[,i] <- lag_wrap(y = fig3a_valence_cor_matrix_permute[,i], n = i)
    fig3a_arousal_cor_matrix_permute[,i] <- lag_wrap(y = fig3a_arousal_cor_matrix_permute[,i], n = i)
    fig3a_spam_cor_matrix_permute[,i] <- lag_wrap(y = fig3a_spam_cor_matrix_permute[,i], n = i)
    fig3a_prob_cor_matrix_permute[,i] <- lag_wrap(y = fig3a_prob_cor_matrix_permute[,i], n = i)
    fig3a_intext_cor_matrix_permute[,i] <- lag_wrap(y = fig3a_intext_cor_matrix_permute[,i], n = i)
  }
  
  #### Test correlation (lower triangle only)
  gv_r <- cor.test(fig3a_gamble_cor_matrix[lower.tri(fig3a_gamble_cor_matrix)], 
                   fig3a_valence_cor_matrix_permute[lower.tri(fig3a_valence_cor_matrix_permute)], method = "spearman", exact = FALSE)
  ga_r <- cor.test(fig3a_gamble_cor_matrix[lower.tri(fig3a_gamble_cor_matrix)], 
                   fig3a_arousal_cor_matrix_permute[lower.tri(fig3a_arousal_cor_matrix_permute)], method = "spearman", exact = FALSE)
  gs_r <- cor.test(fig3a_gamble_cor_matrix[lower.tri(fig3a_gamble_cor_matrix)], 
                   fig3a_spam_cor_matrix_permute[lower.tri(fig3a_spam_cor_matrix_permute)], method = "spearman", exact = FALSE)
  gi_r <- cor.test(fig3a_gamble_cor_matrix[lower.tri(fig3a_gamble_cor_matrix)], 
                   fig3a_intext_cor_matrix_permute[lower.tri(fig3a_intext_cor_matrix_permute)], method = "spearman", exact = FALSE)
  gp_r <- cor.test(fig3a_gamble_cor_matrix[lower.tri(fig3a_gamble_cor_matrix)], 
                   fig3a_prob_cor_matrix_permute[lower.tri(fig3a_prob_cor_matrix_permute)], method = "spearman", exact = FALSE)
  
  #### Store
  gamble_valence_r$r[j] <- as.numeric(gv_r$estimate)
  gamble_valence_r$p[j] <- as.numeric(gv_r$p.value)
  gamble_arousal_r$r[j] <- as.numeric(ga_r$estimate)
  gamble_arousal_r$p[j] <- as.numeric(ga_r$p.value)
  gamble_spam_r$r[j] <- as.numeric(gs_r$estimate)
  gamble_spam_r$p[j] <- as.numeric(gs_r$p.value)
  gamble_intext_r$r[j] <- as.numeric(gi_r$estimate)
  gamble_intext_r$p[j] <- as.numeric(gi_r$p.value)
  gamble_prob_r$r[j] <- as.numeric(gp_r$estimate)
  gamble_prob_r$p[j] <- as.numeric(gp_r$p.value)
}

# Save for future
save(gamble_valence_r, file = str_c(dir_clean, "/bootstrap/gamble_valence_bootstrap.Rdata"))
save(gamble_arousal_r, file = str_c(dir_clean, "/bootstrap/gamble_arousal_bootstrap.Rdata"))
save(gamble_spam_r, file = str_c(dir_clean, "/bootstrap/gamble_spam_bootstrap.Rdata"))
save(gamble_intext_r, file = str_c(dir_clean, "/bootstrap/gamble_intext_bootstrap.Rdata"))
save(gamble_prob_r, file = str_c(dir_clean, "/bootstrap/gamble_prob_bootstrap.Rdata"))

##### Figure 3a - Cor of gamble matrix with spam matrix ######
fig3a_gs_cor <- cor.test(fig3a_gamble_cor_matrix[lower.tri(fig3a_gamble_cor_matrix)], 
                         fig3a_spam_cor_matrix[lower.tri(fig3a_spam_cor_matrix)], method = "spearman", exact = FALSE)

quantile(gamble_spam_r$r, .99) # compare against bootstrap
quantile(gamble_valence_r$r, .99) # compare against bootstrap

# fig3a - Permutation with true value
fig3a_plot <- ggplot(gamble_spam_r, aes(x = r)) + 
  geom_histogram(fill = "white", color = "black") +
  geom_vline(aes(xintercept = fig3a_gs_cor$estimate), linetype = "dashed", color = "red") + 
  theme_classic() + 
  theme(text = element_text(size = 12))
fig3a_plot
ggsave(filename=str_c(dir_graphs, "/fig3a.pdf"), fig3a_plot , width = 4, height = 4, useDingbats=F)
```

# Figure 4a - Conceptual representations of uncertainty predict risky decision-making

A) The correlation coefficients matrix for all tasks. Results reveal that gambling behavior is predicted by conceptual representations of uncertainty, probability, and valence.

```{r fig4a}
# Correlation plot of all measures with spam 
## Figure 4 - Correlation of gambling behavior and of spam (and all others)

#### Gamble matrix ####
fig4a_gamble <- df_gambles %>%
  filter(sub != 59, sub != 115) %>% # two have to be excluded for 0 variance in one of the other tasks, making correlation plot impossible
  mutate(choice = case_when(choice == 2 ~ 1, 
                            choice == 1 ~ 0)) %>%
  arrange(sub, gamble)

n_subs <- fig4a_gamble %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig4a_gamble %>% select(sub) %>% unique() %>% pull()
n_gambles <- fig4a_gamble %>% select(gamble) %>% unique() %>% tally() %>% pull()

fig4a_gamble_matrix <- matrix(1, n_gambles, n_subs)

for (i in seq(1:n_subs)) {
  fig4a_gamble_matrix[,i] <- fig4a_gamble %>% filter(sub == sub_nums[i]) %>% pull(choice)
}

#### Valence & arousal matrices ####
fig4a_darm <- df_darm %>%
  filter(sub != 59, sub != 115) %>%
  arrange(sub, uncertain_word)

n_subs <- fig4a_darm %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig4a_darm %>% select(sub) %>% unique() %>% pull()
n_words <- fig4a_darm %>% select(uncertain_word) %>% unique() %>% tally() %>% pull()

fig4a_valence_matrix <- matrix(1, n_words, n_subs)
fig4a_arousal_matrix <- matrix(1, n_words, n_subs)

for (i in seq(1:n_subs)) {
  fig4a_valence_matrix[,i] <- fig4a_darm %>% filter(sub == sub_nums[i]) %>% pull(valence)
  fig4a_arousal_matrix[,i] <- fig4a_darm %>% filter(sub == sub_nums[i]) %>% pull(arousal)
}

#### Probability Matrix ####
fig4a_prob <- df_prob %>%
  filter(sub != 59, sub != 115) %>%
  arrange(sub, uncertain_word)

n_subs <- fig4a_prob %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig4a_prob %>% select(sub) %>% unique() %>% pull()
n_words <- fig4a_prob %>% select(uncertain_word) %>% unique() %>% tally() %>% pull()

fig4a_prob_matrix <- matrix(1, n_words, n_subs)

for (i in seq(1:n_subs)) {
  fig4a_prob_matrix[,i] <- fig4a_prob %>% filter(sub == sub_nums[i]) %>% pull(probability)
}

#### Internal external #### Matrix
fig4a_intext <- df_internal_external %>%
  filter(sub != 59, sub != 115) %>%
  arrange(sub, uncertain_word)

n_subs <- fig4a_intext %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig4a_intext %>% select(sub) %>% unique() %>% pull()
n_words <- fig4a_intext %>% select(uncertain_word) %>% unique() %>% tally() %>% pull()

fig4a_intext_matrix <- matrix(1, n_words, n_subs)

for (i in seq(1:n_subs)) {
  fig4a_intext_matrix[,i] <- fig4a_intext %>% filter(sub == sub_nums[i]) %>% pull(internal_external)
}

#### SPAM #### Matrix
fig4a_spam <- df_spam %>%
  filter(sub != 59, sub != 115) %>%
  arrange(sub, uncertain1, uncertain2)

n_subs <- fig4a_spam %>% select(sub) %>% unique() %>% tally() %>% pull()
sub_nums <- fig4a_spam %>% select(sub) %>% unique() %>% pull()
n_words <- fig4a_spam %>% select(uncertain1, uncertain2) %>% unique() %>% tally() %>% pull()

fig4a_spam_matrix <- matrix(1, n_words, n_subs)

for (i in seq(1:n_subs)) {
  fig4a_spam_matrix[,i] <- fig4a_spam %>% filter(sub == sub_nums[i]) %>% pull(rdm_values)
}

#### Correlation matrix: gamble x valence
fig4a_gamble_cor_matrix <- cor(fig4a_gamble_matrix, method = "spearman")
fig4a_valence_cor_matrix <- cor(fig4a_valence_matrix, method = "spearman")
fig4a_arousal_cor_matrix <- cor(fig4a_arousal_matrix, method = "spearman")
fig4a_prob_cor_matrix <- cor(fig4a_prob_matrix, method = "spearman")
fig4a_intext_cor_matrix <- cor(fig4a_intext_matrix, method = "spearman")
fig4a_spam_cor_matrix <- cor(fig4a_spam_matrix, method = "spearman")

#### Correlation of correlation matrices (only keeping lower triangle)
fig4a_gv <- cor.test(fig4a_gamble_cor_matrix[lower.tri(fig4a_gamble_cor_matrix)], 
                     fig4a_valence_cor_matrix[lower.tri(fig4a_valence_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_ga <- cor.test(fig4a_gamble_cor_matrix[lower.tri(fig4a_gamble_cor_matrix)],
                     fig4a_arousal_cor_matrix[lower.tri(fig4a_arousal_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_gs <- cor.test(fig4a_gamble_cor_matrix[lower.tri(fig4a_gamble_cor_matrix)],
                     fig4a_spam_cor_matrix[lower.tri(fig4a_spam_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_gp <- cor.test(fig4a_gamble_cor_matrix[lower.tri(fig4a_gamble_cor_matrix)],
                     fig4a_prob_cor_matrix[lower.tri(fig4a_prob_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_gi <- cor.test(fig4a_gamble_cor_matrix[lower.tri(fig4a_gamble_cor_matrix)],
                     fig4a_intext_cor_matrix[lower.tri(fig4a_intext_cor_matrix)], method = "spearman", exact = FALSE)

fig4a_va <- cor.test(fig4a_valence_cor_matrix[lower.tri(fig4a_valence_cor_matrix)], 
                     fig4a_arousal_cor_matrix[lower.tri(fig4a_arousal_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_vs <- cor.test(fig4a_valence_cor_matrix[lower.tri(fig4a_valence_cor_matrix)],
                     fig4a_spam_cor_matrix[lower.tri(fig4a_spam_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_vp <- cor.test(fig4a_valence_cor_matrix[lower.tri(fig4a_valence_cor_matrix)],
                     fig4a_prob_cor_matrix[lower.tri(fig4a_prob_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_vi <- cor.test(fig4a_valence_cor_matrix[lower.tri(fig4a_valence_cor_matrix)],
                     fig4a_intext_cor_matrix[lower.tri(fig4a_intext_cor_matrix)], method = "spearman", exact = FALSE)

fig4a_as <- cor.test(fig4a_arousal_cor_matrix[lower.tri(fig4a_arousal_cor_matrix)],
                     fig4a_spam_cor_matrix[lower.tri(fig4a_spam_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_ap <- cor.test(fig4a_arousal_cor_matrix[lower.tri((fig4a_arousal_cor_matrix))], 
                     fig4a_prob_cor_matrix[lower.tri(fig4a_prob_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_ai <- cor.test(fig4a_arousal_cor_matrix[lower.tri(fig4a_arousal_cor_matrix)], 
                     fig4a_intext_cor_matrix[lower.tri(fig4a_intext_cor_matrix)], method = "spearman", exact = FALSE)

fig4a_sp <- cor.test(fig4a_spam_cor_matrix[lower.tri(fig4a_spam_cor_matrix)], 
                     fig4a_prob_cor_matrix[lower.tri(fig4a_prob_cor_matrix)], method = "spearman", exact = FALSE)
fig4a_si <- cor.test(fig4a_spam_cor_matrix[lower.tri(fig4a_spam_cor_matrix)], 
                     fig4a_intext_cor_matrix[lower.tri(fig4a_intext_cor_matrix)], method = "spearman", exact = FALSE)

fig4a_pi <- cor.test(fig4a_prob_cor_matrix[lower.tri(fig4a_prob_cor_matrix)], 
                     fig4a_intext_cor_matrix[lower.tri(fig4a_intext_cor_matrix)], method = "spearman", exact = FALSE)

## Cor plot
fig4a_data <- matrix(nrow = 6, ncol = 6)
rownames(fig4a_data) <- c("gamble", "spam", "probability", "valence", "arousal", "source")
colnames(fig4a_data) <- c("gamble", "spam", "probability", "valence", "arousal", "source")

fig4a_data[,1] <- c(1, fig4a_gs$estimate, fig4a_gp$estimate, fig4a_gv$estimate, fig4a_ga$estimate, fig4a_gi$estimate)
fig4a_data[,2] <- c(fig4a_gs$estimate, 1, fig4a_sp$estimate, fig4a_vs$estimate, fig4a_as$estimate, fig4a_si$estimate)
fig4a_data[,3] <- c(fig4a_gp$estimate, fig4a_sp$estimate, 1, fig4a_vp$estimate, fig4a_ap$estimate, fig4a_pi$estimate)
fig4a_data[,4] <- c(fig4a_gv$estimate, fig4a_vs$estimate, fig4a_vp$estimate, 1, fig4a_va$estimate, fig4a_vi$estimate)
fig4a_data[,5] <- c(fig4a_ga$estimate, fig4a_as$estimate, fig4a_ap$estimate, fig4a_va$estimate, 1, fig4a_ai$estimate)
fig4a_data[,6] <- c(fig4a_gi$estimate, fig4a_si$estimate, fig4a_pi$estimate, fig4a_vi$estimate, fig4a_ai$estimate, 1)

pdf(file = str_c(dir_graphs, "/fig4a.pdf"))
corrplot(fig4a_data, col.lim = c(0,1),  addCoef.col ='black', diag = FALSE, type = "upper")
dev.off()
```

# Figure 4b

Figure 4b in the manuscript.

```{r fig4b}
# Average spam participant correlation to create an "average similarity" of SpAM behavior to everyone else
# This requires removing the identity and averaging over the column

fig4a_spam_cor_matrix_copy <- fig4a_spam_cor_matrix
diag(fig4a_spam_cor_matrix_copy) <- NA # remove diagonal (all 1s)
spam_cor_similarity <- data.frame(spam_avg_cor = fig4a_spam_cor_matrix_copy %>%
                                    colMeans(na.rm = T)) %>%
  mutate(sub = fig4a_spam %>% select(sub) %>% unique() %>% pull(), 
         median_split = case_when(spam_avg_cor > median(spam_avg_cor) ~ "similar", 
                                  spam_avg_cor <= median(spam_avg_cor) ~ "dissimilar"))

# Average valence and probability ratings based on median split for spam similarity 
fig4b_data <- spam_cor_similarity %>%
  left_join(., df_darm %>% select(sub, uncertain_word, valence), by = "sub") %>%
  left_join(., df_prob %>% select(sub, uncertain_word, probability), by = c("sub", "uncertain_word")) %>%
  mutate(certain_group = case_when(uncertain_word == "risky" ~ "uncertain", 
                                   uncertain_word == "erratic" ~ "uncertain", 
                                   uncertain_word == "improbable" ~ "uncertain", 
                                   uncertain_word == "unexpected" ~ "uncertain", 
                                   uncertain_word == "doubtful" ~ "uncertain", 
                                   uncertain_word == "uncertain" ~ "uncertain",
                                   uncertain_word == "questionable" ~ "uncertain", 
                                   uncertain_word == "unknown" ~ "uncertain", 
                                   uncertain_word == "ambiguous" ~ "uncertain", 
                                   uncertain_word == "vague" ~ "uncertain", 
                                   uncertain_word == "frequent" ~ "certain", 
                                   uncertain_word == "likely" ~ "certain", 
                                   uncertain_word == "possible" ~ "certain", 
                                   uncertain_word == "predictable" ~ "certain",
                                   uncertain_word == "confident" ~ "certain", 
                                   uncertain_word == "certain" ~ "certain", 
                                   uncertain_word == "definite" ~ "certain", 
                                   uncertain_word == "inevitable" ~ "certain"))

fig4b_sub_data <- fig4b_data %>% 
  group_by(median_split, sub, spam_avg_cor, certain_group) %>%
  summarise(mean_sub_valence = mean(valence), mean_sub_prob = mean(probability), 
            sd_sub_valence = sd(valence), sd_sub_prob = sd(probability))

# Plot
fig4b_avg_data <- fig4b_sub_data %>%
  group_by(median_split, certain_group) %>%
  summarise(mean_valence = mean(mean_sub_valence), sd_valence = sd(mean_sub_valence),
            n_valence = n(), se_valence = sd_valence / sqrt(n_valence), 
            mean_prob = mean(mean_sub_prob), sd_prob = sd(mean_sub_prob),
            n_prob = n(), se_prob = sd_prob / sqrt(n_prob))

group_colors <- c("#0175BB", "#F0C11A")

fig4b_plot1 <- ggplot(fig4b_avg_data, aes(x = median_split, y = mean_valence, fill = certain_group)) + 
  geom_col(position = position_dodge(.9)) + 
  geom_errorbar(aes(ymin = mean_valence - se_valence, ymax = mean_valence + se_valence), 
                position = position_dodge(.9), width = .1) + 
  scale_y_continuous(name = "Valence") + 
  scale_fill_manual(values = group_colors) + 
  xlab("SPaM Group") + 
  theme_classic()

fig4b_plot2 <- ggplot(fig4b_avg_data, aes(x = median_split, y = mean_prob, fill = certain_group)) + 
  geom_col(position = position_dodge(.9)) + 
  geom_errorbar(aes(ymin = mean_prob - se_prob, ymax = mean_prob + se_prob), 
                position = position_dodge(.9), width = .1) + 
  scale_y_continuous(name = "Probability") + 
  scale_fill_manual(values = group_colors) + 
  xlab("SPaM Group") + 
  theme_classic()

fig4b_plot <- cowplot::plot_grid(fig4b_plot1, fig4b_plot2, labels = c("A", "B"))
fig4b_plot

ggsave(filename = str_c(dir_graphs, "/fig4b.pdf"), fig4b_plot, width = 6, height = 4)

# Regression stats 
fig4b_model1 <- lmer(valence ~ median_split*certain_group + (1 + certain_group | sub), 
                       data = fig4b_data)
summary(fig4b_model1)

fig4b_model2 <- lmer(probability ~ median_split*certain_group + (1 + certain_group | sub), 
                       data = fig4b_data)
summary(fig4b_model2)

# T-test on difference (valence)
ttest_1 <- fig4b_data %>% 
  group_by(sub, certain_group, median_split) %>%
  summarise(mean_sub_valence = mean(valence)) %>%
  pivot_wider(names_from = certain_group, values_from = "mean_sub_valence") %>%
  mutate(valence_diff = uncertain - certain)

t.test(ttest_1$valence_diff[ttest_1$median_split == "dissimilar"], 
       ttest_1$valence_diff[ttest_1$median_split == "similar"])

# T-test on difference (probability)
ttest_2 <- fig4b_data %>% 
  group_by(sub, certain_group, median_split) %>%
  summarise(mean_sub_prob = mean(probability)) %>%
  pivot_wider(names_from = certain_group, values_from = "mean_sub_prob") %>%
  mutate(prob_diff = uncertain - certain)

t.test(ttest_2$prob_diff[ttest_2$median_split == "dissimilar"], 
       ttest_2$prob_diff[ttest_2$median_split == "similar"])
```